% $Header: /home/mwicks/Projects/Gaspra-projects/cvs2darcs/Repository-for-sourceforge/dvipdfm-initial/doc/dvipdfm.tex,v 1.13 1998/11/21 20:18:57 mwicks Exp $
\special{  pdf:  docview <</PageMode /UseOutlines>>}%
\catcode`\ =11\def\space{ }\catcode`\ =10
% Page layout
\magnification\magstephalf
\advance\hoffset by 0.5truein
\advance\voffset by 0.5truein
\advance\hsize by -1.0in
\advance\vsize by -1.0in
\def\makefootline{\baselineskip=40pt\line{\the\footline}}
%
%  Some helpful symbols
\def\rtm{{\font\r=cmss10 at 4pt
\font\c=cmsy5
\setbox0\hbox{\c\char13}\skip0\wd0\box0\setbox0\hbox{\r R}\advance\skip0 by
\wd0\kern-0.5\skip0\box0}}
\def\tm{{\font\r=cmss10 at 4pt \hbox{\r TM}}}%
%
%
% Some color definitions
% 
\def\begincolor#1{\special{pdf:bc #1}}%
\def\endcolor{\special{pdf:ec}}%
\def\colored#1#2{%
  \begincolor{#1}#2\endcolor}
\def\red{[0.8 0.5 0]}%
\def\green{[0 1 0]}%
\def\blue{[0 0.4 0.8]}%
\def\yellow{[0.8 0.9 0.1]}%
\def\linkcolor{[0.5 0.0 0.7]}%
%
% Small help for images
\def\reserve#1#2#3{\setbox0\hbox{}\wd0=#1
                   \ht0=#2\special{#3}\box0}
\def\nreserve#1#2#3{\setbox0\hbox{}\wd0=#1
                   \ht0=#2{#3}\box0}
%
% Some font definitions
%
\font\maintitlefont=cmti12 at 20.74pt
\font\headingfont=cmss12 at 14.4pt
\font\subheadingfont=cmss12 at 12pt
%
%
% Miscellaneous token lists
%
\newtoks\title\newtoks\author\newtoks\version
\newtoks\date
%
% Bibliography counters
\newcount\biblioitems\biblioitems=0
\def\bibitem{\medskip\advance\biblioitems by 1\item{[\the\biblioitems]}}
%
% Counters for section levels
%
\newcount\sectioncount\newcount\ssectioncount\newcount\sssectioncount
\sectioncount0\ssectioncount0\sssectioncount0
%
\newskip\indentlevel\indentlevel=\parindent
\newskip\secskip\secskip=24pt plus 1pt minus 2pt
\newskip\subsecskip\subsecskip=16pt plus 1pt minus 2pt
\def\beginlist{\par\advance\leftskip by \indentlevel\advance\rightskip by
\indentlevel\medskip}
\def\endlist{\par\advance\leftskip by -\indentlevel\advance\rightskip by
-\indentlevel\medskip}
%
%
\def\settitle{%
  {\maintitlefont\colored{\blue}{\the\title}}%
  \special {pdf:  docinfo << /Title (\expandafter\the\title) >>}}%
%
\def\setauthor{%
  {\headingfont\colored{\blue}{\the\author}}%
  \special {pdf:  docinfo << /Author (\the\author) >>}}%
%
\def\setversion{%
  {\subheadingfont{\the\version}}%
}%
\def\setdate{%
  {\subheadingfont{\the\date}}%
}%
\def\setlink#1{\colored{\linkcolor}{#1}}%
%  
\def\setheading#1{%
  {\headingfont\colored{\blue}{#1}}\raise\baselineskip\hbox{\special{pdf: outline 1 << /Title (#1) /Dest [
@thispage /FitH @ypos ]  >> }}}%
%
\def\setsubheading#1{%
  {\subheadingfont\colored{\blue}{#1}}\raise\baselineskip\hbox{\special{pdf: outline 2 << /Title (#1) /Dest [
@thispage /FitH @ypos ]  >> }}}%
%
\def\link#1#2{\setbox0\hbox{\setlink{#1}}%
   \special{pdf: ann width \the\wd0\space height \the\ht0\space depth \the\dp0
     << /Type /Annot /Subtype /Link /Border [ 0 0 0 ] /A << /S /GoTo
     /D (#2) >> >>}\box0\relax}%
%
\def\dest#1{\hbox{\raise14pt\hbox{\special{pdf:dest (#1) [ @thispage /FitH
@ypos ]}}}}%
%
\def\maketitle{\noindent\settitle\hfill\setauthor\par
          \hrule height 1.0pt\medskip
          \noindent\setversion\hfill\setdate\vskip0.3in}%
%
\def\section#1{\advance\sectioncount by 1\ssectioncount0
\vskip-\lastskip\vskip\secskip\goodbreak
\noindent\setheading{\the\sectioncount. #1}\medskip\nobreak}%
\def\subsection#1{\vskip-\lastskip\advance\ssectioncount by 1%
\vskip\subsecskip\noindent
\setsubheading{\the\sectioncount.\the\ssectioncount\
          #1}\par\nobreak\smallskip}%
%
\def\display#1{\medskip\line{\quad #1\hfil}\medskip}
\def\example{\medskip\noindent{\it Example:\quad\nobreak}}
\def\note{\medskip\noindent{\it Note:\quad\nobreak}}
\def\syntax{\medskip\noindent{\it Syntax:\quad}}
\def\description{\medskip\noindent{\it Description:\quad}}
%
\def\newpage{\vfill\eject}
%
\def\dvipdfm{{\tt dvipdfm}}%
\input verbatim
%
\title{Dvipdfm User's Manual}
\author{Mark A. Wicks}
\version{Version 0.7dev}
\date{November, 1998}
\maketitle
\section{Introduction}
\special{pdf:epdf (xp56ref.pdf)}
This package is a DVI (\TeX) to PDF conversion utility,
having the following features:

\special{pdf: epdf (5004.AFM_Spec.pdf) }

\beginlist
\item{$\bullet$} Support for outline (bookmark) entries, named destinations,
and \link{annotations}{annotate} (including hyperlinks, forms and widgets).  Nearly
every Acrobat Distiller pdfmark is approximated.

\item{$\bullet$} Support for arbitrary, nested linear transformations
of typeset material.  Any material on the page, including
\TeX\ text, may be \link{scaled}{textxform} \link{and}{textxform} \link{rotated}{textxform}.

\item{$\bullet$} Ability to include a \link{PDF file}{pdfimage} as
an encapsulated object along with its embedded resources such as fonts.
Note:  Currently, this doesn't work for linearized PDF files or
second generation PDF files.

\item{$\bullet$} Ability to include a \link{JPEG image}{jpegimage} as an encapsulated
object.

\item{$\bullet$} A color stack.
\endlist

The electronic version of the document exercises
some of the hypertext features and serves as
a sample input file for |dvipdfm|.
It assumes the reader has some familiarity with the basic features
of the Portable Document Format.  The PDF specification
is distributed by Adobe Systems\link{[1]}{pdfmanual}.
An excellent source for information about PDF documents
in general is \link{[2]}{pdfprimer}.  Information about
using \TeX\ to construct PDF documents (mainly using Distiller) is the
AcroTeX home page\link{[3]}{acrotex}.

Currently, the widely accepted method to generate PDF file from \TeX\
is to use Adobe's Acrobat Distiller on a PostScript
file produced by {\tt dvips}.
The hyperlink features are accessed by using \TeX\ |special| primitives
to embed pdfmarks in the PostScript produced by |dvips|.
H\`an Th\'e Th\`an's PDF\TeX\ project is an alternative method
of generating PDF from \TeX\ source.
Although quite good and fairly mature, the PDF\TeX\ project
modified \TeX\ itself to add primitives that support the PDF features.
I prefer to work with \TeX\ unmodified, as released by Donald Knuth
(call me a purist).
There is an existing DVI to PDF driver called
|dvipdf| written by Sergey Lesenko.  At present, it's not widely
available, so I haven't used it.  I wrote |dvipdfm|
mainly as an exercise to get at the features
of PDF I was trying to use.  This |dvipdfm| project demonstrates that many features
of PDF can be accessed by using a DVI driver.
The PDF features are activated in the driver via
\TeX\ |special| primitives.

Even though Distiller is the best method of generating PDF (and
probably will remain so for some time) I have several reasons for
seeking alternatives to Distiller.
First, Distiller isn't available for my principle operating
system---Linux.

My second objection is philosophical.
\TeX\ is a programming language.
A DVI file is a page description consisting of very
simple program instructions that have no branching or
decision instructions.
Similarly PostScript is a complete programming language,
while PDF is a page description language consisting
of simple program instructions
without any branching or decision capabilities.
\TeX\ is like PostScript (without the graphics)
while DVI is like PDF (without the graphics or the hyperlinks).
Creating PDF from DVI using Distiller requires converting a page description to a program,
and converting that program back to a page description.
To continue this analogy,
Pdfmarks are PostScript ``escapes'' and are meant for the Distiller.
\TeX\ |\special| primitives are \TeX\ ``escapes'' and are meant for the DVI driver.
It seems natural to go directly from DVI to PDF, where \TeX\ replaces
PostScript, the DVI driver replaces Distiller,
and \TeX\ |\special| primitives replace the pdfmarks.

Unfortunately, until graphics software
begins to produce PDF content streams or encapsulated
PDF objects, PostScript will remain the easiest
way to include graphics in \TeX\ documents.
I would hope that in the future, graphics programs
will begin to produce PDF content streams or PDF objects that
may be included using a DVI to PDF translator.  Either
of these may be easily embedded using \dvipdfm\ or a similar driver.

\section{General Concepts and Syntax}

Each \TeX\ |\special|
represents a separate command to the \dvipdfm\ 
driver.  Each |special| must begin with ``|pdf:|''
to identify that ||special|| as a command for the \dvipdfm\ driver.
A |\special| beginning with any other characters is ignored
by the driver.  Leading spaces are ignored.  The characters ``|pdf:|''
are immediately followed by a \dvipdfm\ command.  These commands
are documented in \hbox{Sections~3--6}.

\subsection{PDF Object Syntax and Variable Expansion}
With one exception, the syntax used for PDF objects
within each |\special| specials follows
the PDF specification.
The one exception is variable expansion.
In the syntax specifications that follow, {\it PDF\_Object}
means that an arbitary PDF object is expected.  Similarly
{\it PDF\_Array} indicates that a PDF array is expected, {\it PDF\_Dict}
inciates that a PDF dictionary is expected, etc.
See the \link{reference manual}{pdfmanual}\
for a complete list of PDF object types.

The single extension implemented in this driver
allows a symbol name of the
form |@|{\it name} whereever any PDF object is expected.
The {\it name} may contain any characters allowed
in a PDF name.
A user-defined symbol beginning with |@| expands to an indirect
reference to the user-defined PDF object.  This feature replaces
the |{|{\it name}|}| syntax used with pdfmarks.
In addition to the user-defined names, some names
are defined by the driver.
The driver defined variables are for
referencing objects such as the current page, future pages,
or the current location on the current page.
The driver defined variables appear in \link{Table~1}{drivervariables}.
\topinsert
{\dest{drivervariables}}
\centerline{\colored{\blue}{\subheadingfont Table~1---List of driver
defined variables}}
\bigskip\nobreak
\centerline{\vbox{\halign{{\tt #}\hskip
1em&\vtop{\leftskip0pt\rightskip0pt\hsize=3.0in\noindent #}\cr
\omit\hfil\it Variable\hfil&\omit\hfil \it Description \hfil\cr
\noalign{\smallskip\begincolor\red\hrule\endcolor\smallskip}
@thispage&An {\it indirect reference} to the current page.\cr 
@page&An {\it indirect reference} to page $n$.\cr
@nextpage&An {\it indirect reference} to the page following the current page.\cr
@prevpage&An {\it indirect reference} to the page preceding the current page.\cr
@ypos&A {\it number} representing the current vertical position in units of PDF points.\cr
@xpos&A {\it number} representing the current horizontal position in units of PDF points.\cr
}}}
\endinsert

In the syntax specifications that follow, several
standard conventions are followed.  Terminal
characters that appear in the command 
are typeset in the |\tt| font, e.g., |object|.
Nonterminal symbols are typeset in italics.
Optional parameters are surrounded by brackets, e.g.,
[{\it optional\_argument}].  An item followed
by~``{*}'' represents an item that may appear
zero or more times.  An item followed by~``{+}''
represents a required item that may appear multiple times.

\subsection{Dimensions and transformations}
Interaction with the |dvipdfm| driver consists
of short commands with a few arguments delimited by white space.
Typically the arguments are PDF objects.
Two exceptions are dimension specifications and transformations.

In the \TeX\ style, a dimension specification consists of one of the keywords
|width|, |height|, or |depth| followed by a dimension
consisting of a numerical value, followed by a unit for the dimension.  The
unit will typically be |pt| (which represents a \TeX\ point, not a
PDf point) but |cm| and |in| are also allowed.  The notation
{\it dimension\/} in a syntax description means a dimension is expected.

A transformation consists of one of the keywords |scale|, |xscale|,
|yscale|, or |rotate| followed by a numerical value.  In the
case of |rotate| the value is the rotation angle in degrees.
The notation
{\it transformation} means a transformation is expected. 

\section{Document Construction Commands}
All commands are executed via \TeX\ |\special| primitives
prefixed with the characters ``|pdf:|''.
\example
\begintt
\special{ pdf: out 1 << /Title (Introduction)
                        /Dest [ 1 0 R /FitH 234 ] >>
\endtt

\dest{annotate}
\subsection{Annotate}
\hbox to 0pt{\hskip-1.0in\special{pdf: ann width 3.0in height 36pt << /Type /Annot /Subtype /Text
/Contents (This is a /Text Annotation.  Aren't these things ugly?.  It's a
good thing they don't print by default.) >>}\hss}
\syntax
{\tt annotate} [{\tt @}{\it name}] {\it dimension}+ {\it PDF\_dictionary}
\description
The |annotate| (|annot| or |ann|) command defines an annotation.
Annotations are typically used for
notes, hyperlinks, forms, or widgets.
The parameter {\it name} is an optional alphanumeric identifier
and {\it PDF\_dictionary} is a valid PDF dictionary after variable expansion.
If {\tt @}{\it name} is specified, it may be used in
other PDF objects to refer to this annotation.
One or more {\it dimension} parameters are required
and each consists of the keyword
{\tt height}, {\tt width}, or {\tt depth} followed
by an appropriate length, specified as per \TeX\null.
The |width| must be nonzero and either the |height| or |depth|
must be nonzero.
Each length is a number followed by a unit, such as {\tt pt},
{\tt in}, or {\tt cm}.  Since these values
would typically be entered by \TeX, a {\tt pt}
is a \TeX\ point, not a PDF point.

\example The annotation in this subsection
was typeset with
\begintt
\special{pdf: ann width 3.0in height 36pt
     << /Type /Annot /Subtype /Text
        /Contents (This is a /Text Annotation.
                   Aren't these things ugly?.
                   It's a good thing they don't
                   print by default.) >>}
\endtt


\subsection{Article}
\syntax
{\tt article} {\tt @}{\it name} {\it PDF\_dictionary}

\description
The |article| (or |art|) command initializes an article.  An article
is a collection of boxed regions in the document that should be
read consecutively. The {\it name} parameter is required.  The required PDF dictionary 
is similar to the |/Info| dictionary accessed via the |docinfo| command
and would typically include the |/Title| and |/Author| keys.
\example
\begintt
\special {pdf: article @somearticle << /Title (Some title)
                                       /Author (Me) >>}
\endtt

\subsection{Bead}
\syntax
{\tt bead} {\tt @}{\it name} {\it dimension}+

\description
The |bead| command adds a rectangular area to an existing article thread.
The parameter {\it dimension}+ specifies a rectangular area
in the same manner as for an annotation.  The {\it name}
must correspond to an existing |article|.
\example
\begintt
\special{pdf: bead @someart width 156pt height 20pt depth 4pt}
\endtt

\subsection{Dest}
\syntax
{\tt dest} {\it PDF\_String} {\it PDF\_Dest}
\description
The |dest| command defines a named destination.
The {\it PDF\_String} is a PDF string naming
the destination.  This string may be used in the destination
fields of annotations and outline entries to refer to
this destination.  {\it PDF\_Dest} is a PDF
destination object (typically an array).
\example
\begintt
\special{pdf: dest (listofreferences) [ @thispage /FitH @ypos ]}
\endtt

\subsection{Docinfo}
\syntax
{\tt docinfo} {\it PDF\_dictionary}
\description
The |docinfo| command adds the keys in the specified dictionary to the
document's |/Info| dictionary.  All keys are optional, but may include
the keys |/Author|, |/Title|, |/Keywords|, |/Subject|,
and |/Creator|.
\example
\begintt
\special{pdf: docinfo << /Author (Mark A. Wicks)
                         /Title (This Document) >>}
\endtt

\subsection{Docview}
\syntax
{\tt docview} {\it PDF\_dictionary}
\description
The |docview| command adds the keys in the specified dictionary to the
document's |/Catalog| dictionary.  All keys are optional, but may include
the keys |/PageMode|,
|/URI|, |/OpenAction|, |/AA|
and |/ViewerPreferences|.  See the PDF Reference Manual
for documentation of these keys and additional keys.
\example
\begintt
\special{pdf: docview << /PageMode /UseThumbs >> }
\endtt


\subsection{Object}
\syntax
{\tt object} [@{\it name}] {\it PDF\_Object}
\description
The |object| (also |obj|) command creates a
PDF object.  The parameter {\it PDF\_Object} is any valid PDF object.  The
parameter {\it name} may be used to provide an indirect reference
to this object within other
objects. It will be expanded anywhere within
a {\tt special} where a PDF object is
expected. Typically {\it object} is an array
or dictionary.  It may be an empty array or
dictionary that can be constructed dynamically via
the {\tt put} command.
\example
\begintt
\special{pdf: object @mydict << /Firstpage @thispage >>}
\endtt

\subsection{Out}
\syntax
{\tt out} {\it number} {\it PDF\_dictionary}

\description
The |out| (also |outline|) command adds an outline (also called a ``bookmark'') entry
to the document.
The parameter {\it level\/} is an integer representing the
level of the outline entry (beginning with 1) and
{\it  PDF\_dictionary\/} must contain
the two keys {\tt /Title} and either {\tt /Dest} or {\tt /A}.
It may also contain the {\tt /AA} key.  These keys are documented
in the PDF Reference Manual.
\example
\begintt
out 1 << /Title (Section 1) /Dest [ @thispage /FitH @ypos ] >>
\endtt
which may be followed by
\begintt
out 2 << /Title (Section 1.1) /Dest [ @thispage /FitH @ypos ] >>
\endtt
\note
You may not skip levels.  A level~2 outline entry
must follow a level~1 outline
entry.  A level~3 outline entry must follow a level~2 outline
and cannot immediately follow a level 1 outline entry.

\subsection{Put}
\syntax
\beginlist
{\tt put} @{\it name} {\it PDF\_Object}
\endlist
or 
\beginlist
{\tt put} @{\it name} {\it PDF\_Dictionary}
\endlist
\description
The |put| command modifies an existing PDF object created with |obj|.
The first form is used when @{\it name} is an array.  The second
form is used when @{\tt name} is a dictionary.  Arrays are
incremented one object at a time.  All keys in {\it PDF\_Dictionary}
are added to the dictionary represented by @{\it name}.
\example
\begintt
\special{pdf: put @mydict << /Nextpage @thispage >>}
\endtt

\subsection{Close}
\syntax
{\tt close} @{\it name}
\description
The |close| writes the named PDF object created with |obj| to the PDF file.
No further |put| commands may be executed for this object.
The object may continue to be referenced using @{\it name}
indefinitely.  If the object is never closed, it will
be closed when |dvipdfm| finishes processing the document.

\section{Text Transformation Commands}
The commands in this section deal with transformation
of arbitrary material, which may include
material typeset by \TeX.  These
may also be used on included graphics images
if the commands in Section~8 won't do the job.


\subsection{BeginTransform}
\syntax
{\tt begintransform} {\it transformation}+

\description
The |begintransform| (|btrans| or |bt|) applies
the specified transformation to all subsequent text.
The scaling is applied first, followed by the rotation.
The reference point of a box following
the |\special| remains fixed.  Such transformations
may be nested to perform rotations within rotated text, for exmaple.

\example
\begintt
\special{pdf: bt rotate 90 xscale 2.0 }
\endtt

\subsection{BeginTransform}
\syntax
{\tt endtransform}

\description
The |endtransform| (|etrans| or |et|) concludes
the action of the immediately preceding |begintransform|
command.  All transformations must be closed
on the same page.  The driver will close
any pending unclosed transformations at the
end of the page and issue a warning message.
All material to be transformed
should probably be enclosed in a single box
to prevent any break.
\example
\begintt
\special{pdf: et}
\endtt


\section{Color Commands}
The commands in this section deal with manipulation of the color
stack.

\subsection{Begincolor}
\syntax
{\tt begincolor} {\it PDF\_Array}
\description
The |begincolor| (|bcolor| or |bc|) command uses the
array to set the default color for future marking operators.
The current color is pushed on the color stack.  The
array must have three elements specifying the coordinates
of the color in the Device RGB color space.
\example
\begintt
\special{ pdf: bc [ 1 0 0 ] }
\endtt

\subsection{Endcolor}
\syntax
{\tt endcolor}
\description
The |endcolor| (|ecolor| or |ec|)
changes the default color to
match the color on the top
of the stack.  It removes
the color from the stack.
\example
\begintt
\special{ pdf: ec }
\endtt


\section{Image Commands}
The commands in this section deal with embedding
graphics into your PDF document.  The present
driver supports PDF and JPEG graphics inclusion.

\subsection{Epdf}
\syntax
{\tt epdf} [|@|{\it name}] [{\it dimension}$\vert${\it transformation}]*  {\it PDF\_String}

\description
The {\tt epdf} command ``encapsulates'' the first page of a PDF
file named by {\it PDF\_String}
into a PDF XObject.  The resulting XObject is drawn
with the lower left corner at the current location of the page.
The optional @{\it name} parameter may be used
to reference this object within other objects.  If a
{\it dimension} is supplied, the object will be scaled to fit
that dimension.  A {\it transformation} consists of one of the keywords
|scale|, |xscale|, |yscale|, or |rotate|
followed by a number representing
the scaling factor or rotation angle in degrees.  Both {\it transformation} and {\it dimension}
parameters can be supplied as long as they are not logically
inconsistent.
\example
\begintt
\special{pdf:epdf yscale 0.50 width 4.0in rotate 45 (circuit.pdf)}
\endtt

\subsection{Image}
\syntax
{\tt image} [@{\it name}] [{\it dimension} $\vert$ {\it transformation}]*  {\it PDF\_String}

\description
The {\tt image} command ``encapsulates'' a JPEG image
taken from the file named by {\it PDF\_String}.
Otherwise, this command functions just like |epdf|.

\section{Raw Page Marking Commands}
The commands in this section deal with embedding
raw PDF graphics operators into your PDF document.

\subsection{Bop}
\syntax
{\tt bop} {\it stream}
\description
The |bop| command specifies a marking
stream to be generated at the top of each page.
The parameter {\it stream} is any sequence
of marking operators and is added to the page's content stream.  
The stream is applied {\it to all pages} regardless
of where it appears in the document.
\example  The two horizontal lines appearing
at the top of each page in this document
were set with
\begintt
\special {pdf: bop  q 0 w 0.8 0.5 0 RG
               54 740 m 504 740 l 504 740.25 l 54 740.25 l b
               36 760 m 504 760 l 504 760.25 l 36 760.25 l b Q }
\endtt
\special {pdf: bop  q 0 w 0.8 0.5 0 RG
                           54 740 m 504 740 l 504 740.25 l 54 740.25 l b  
                           36 760 m 504 760 l 504 760.25 l 36 760.25 l b Q }

\subsection{Content}
\syntax
{\tt content} {\it stream}
\description
The |content| command specifies a marking
stream to be added to the current page at
the current location.  While it
is possible to change the color
state, etc., with this command, it is
not advised.  Use the color management
commands to change colors.


\subsection{Eop}
\syntax
\beginlist
{\tt eop} {\it stream}
\endlist
\description
The {\tt eop} specifies a marking stream to be generated at the end
of each page. The parameter {\it stream} is any sequence
of marking operators and is added to the page's content stream.  
The stream is applied {\it to all pages} regardless
of where it appears in the document.

\section{Graphics Examples}

The examples in this section illustrate some of the transformation
and image inclusion capabilities of \dvipdfm.

\dest{textxform}
\subsection{Text Transformation}
Tables with slanted entries are possible as shown
in \link{Table~2}{rotatedtable}.  This table was achieved using
various ``|bt rotate 35|'' commands.  It
is difficult to do without macro support.

The following line of text was done with nested
combinations of ``|bt rotate 10|'' and ``|bt rotate -10|''.

\def\up#1{\special{pdf:bt rotate 10}\hbox{#1}}
\def\down#1{\special{pdf:bt rotate -10}\hbox{#1}}
You \down{can} \down{nest} \down{the} \up{text}
\up{trans}\up{form}\up{ation} \up{capa}\down{bili}\down{ties} \down{to}
\down{achieve} \down{effects} \down{like}
{this}.
\special{pdf:et}\special{pdf:et}\special{pdf:et}\special{pdf:et}
\special{pdf:et}\special{pdf:et}\special{pdf:et}\special{pdf:et}
\special{pdf:et}\special{pdf:et}\special{pdf:et}
\special{pdf:et}
\special{pdf:et}
\special{pdf:et}
\vskip 0.5in


\def\rvr{\colored{\red}{\vrule}}\def\entry#1{\special{pdf:bt rotate 35}%
{\vbox to 0pt{\vss\hbox{\quad\colored{\blue}{#1}\quad}\vskip 1pt\colored{\red}{\hrule}}\special{pdf:et}}}
\topinsert
\dest{rotatedtable}
\bigskip
\centerline{\colored{\blue}{\subheadingfont Table 2---Example of rotated
text set in Computer Modern Roman}}
\bigskip
\centerline{\vbox to 0.45in{\vss\offinterlineskip\halign{#&#&#&#&#&#\cr
\entry{1994}&
\entry{1995}&
\entry{1996}&
\entry{1997}&
\entry{1998}&
\entry{1999}\cr
\noalign{\colored{\red}{\hrule}}
\strut\rvr&\rvr&\rvr&\rvr&\rvr&\rvr\crcr}}}
\bigskip
\endinsert

\dest{imageinclude}
\subsection{Image Inclusion}
 
The image in \link{Figure~1}{jpegimage} was included from a JPEG file.
The image shown in \link{Figure~2}{rotatedimage} comes from the same file,
but is loaded at a 50\% scale and a 45$^\circ$ rotation.

\topinsert
\dest{jpegimage}
\bigskip
\centerline{\reserve{1.50in}{2.05in}{pdf: image (mwicks.jpeg)}}
\medskip
\centerline{\subheadingfont
\begincolor{\blue}Figure~1---A JPEG image of the author.\endcolor}
\bigskip
\endinsert


\topinsert
\dest{rotatedimage}
\bigskip
\centerline{\reserve{0.75in}{1.25in}{pdf: image rotate 45 scale 0.50 (mwicks.jpeg)}}
\medskip
\centerline{\subheadingfont
\begincolor{\blue}Figure~2---Image of the author scaled by 0.5 and rotated
by 45$^\circ$.%
\endcolor}
\bigskip
\endinsert

Sophisticated macro support for images is not yet available.  For now,
you need to enclose the image
in your own box of the correct size so \TeX\ reserves
space for it. No space is reserved for a |special|
unless you reserve it.  I used the following simple
Macro for the previous images:
\begintt
\def\reserve#1#2#3{\setbox0\hbox{}\wd0=#1
                   \ht0=#2\special{#3}\box0}
\endtt
The arguments to the |\reserve| macro are the width, height,
and |\special| contents, respectively.
The first image in this section was included with
\begintt
\centerline{\reserve{1.50in}{2.05in}{pdf: image (mwicks.jpeg)}}
\endtt

Until macro support is more developed, you will need to know
the dimensions of the image.  By default, JPEG files are included
at a resolution of 100dpi so if you know the pixel size
of the image, you know how much space to reserve.
Any \TeX\ magnification is applied to the image
in addition to any scaling defined in the |\special|.
For example, this document sets |\magnification=\magstephalf|,
so the images are actually scaled by 1.095.  The first
image in this section has a printed width of 1.643in
even though 1.50in was specified in the |\special|.

Several command line utilities are available to read
the pixel dimensions of a JPEG file.  For PDF files,
you can |grep| on |/MediaBox| to get an indication
of the image size.  The |/MediaBox| dimensions are
in PDF points.

\link{Figure~3}{pdfimage} was produced by embedding a PDF file
using |epdf|.

\topinsert
\dest{pdfimage}
\centerline{\reserve{2.41in}{2.6in}{pdf: epdf (transistor.pdf)}}
\medskip\centerline{\subheadingfont
\begincolor{\blue}Figure~3---An embedded PDF object.
\endcolor}
\endinsert
Notice that any resources required for the object
are also embedded.  In this case, the Times Roman font
resource was embedded along with the content stream.

\newpage
\section{References}
\tolerance=1000\hbadness=1000
\bibitem\dest{pdfmanual}{\it Portable Document
Format Reference Manual}, Version 1.2, Adobe Systems Incorporated, 1996.
Available at the following URL:  {\tt http://www.adobe.com}. 

\bibitem\dest{pdfprimer}Thomas Merz,
{\it Web Publishing with Acrobat/PDF},
Springer-Verlag, 1997, ISBN 3-540-63762-1.  Chapter~6 of this book is available
at the URL:  |http://http://www.ifconnection.de/~tm|.

\bibitem\dest{acrotex}D. P. Story, {\it AcroTeX},  The AcroTeX home page
is located at the URL:  |http://www.math.uakron.edu/~dpstory/acrotex.html|.

\bye
